Bioverse: Origins of Life

### 2023-08-16 src/tex/ms.tex
model P_env and show overlaps in plot

 \subsection{Required environment $\pdf_\mathrm{env, i}$}
% Following previous work~\citep{Spiegel2012,Chen2018,Kipping2021}, we may adopt a uniform rate model for abiogenesis, i.e., assume that OoL events occur at a uniform rate.
% This corresponds to a Poisson process with a rate parameter $lambda$, where we make the implicit assumption that abiogenesis occurs only via a single, unique mechanism, only once, and instantaneous.
% If this event occurs within a limited time window $t$, say, between planets form around a star and when it leaves the main sequence, we have
% \begin{align}
% \pdf_\mathrm{env, i} = 1 - \exp(-\lambda t).
% \end{align}
% ~\\
\todo[inline]{The planet should be in the liquid-water HZ}
...
Let us assume that $H_1$ only requires a minimum bulk density $\rho_1$, such that $\pdf_\mathrm{env, 1} = 1$ if $\rho > \rho_1$ and zero otherwise.
On the other hand, $H_2$ requires exposed land and thus a small water mass fraction.
We may implement this in the same way as above but with a minimum bulk density $\rho_2 > \rho_1$.
Furthermore, there is a requirement that the tidal locking timescale may not be so small that the planet is likely tidally locked at observation.
This translates to imposing a minimum semimajor axis $a_2$.
...
...

>>>>

 \subsection{Required environment $\pdf_\mathrm{env, i}$}
% Following previous work~\citep{Spiegel2012,Chen2018,Kipping2021}, we may adopt a uniform rate model for abiogenesis, i.e., assume that OoL events occur at a uniform rate.
% This corresponds to a Poisson process with a rate parameter $lambda$, where we make the implicit assumption that abiogenesis occurs only via a single, unique mechanism, only once, and instantaneous.
% If this event occurs within a limited time window $t$, say, between planets form around a star and when it leaves the main sequence, we have
% \begin{align}
% \pdf_\mathrm{env, i} = 1 - \exp(-\lambda t).
% \end{align}
% ~\\
\todo[inline]{The planet should be in the liquid-water HZ}
...
Let us assume that $H_1$ only requires a minimum bulk density $\rho_1$, such that $\pdf_\mathrm{env, 1} \rightarrow 1$ for $\rho >> \rho_1$ and $\pdf_\mathrm{env, 1} \rightarrow 0$ for $\rho << \rho_1$.
On the other hand, $H_2$ requires exposed land and thus a small water mass fraction.
We may implement this in the same way as above but with a minimum bulk density $\rho_2 > \rho_1$.
Furthermore, there is a requirement that the tidal locking timescale may not be so small that the planet is likely tidally locked at observation.
This translates to imposing a minimum semimajor axis $a_2$.

To approximate these thresholds including their expected intrinsic fuzziness, we model them with logistic sigmoid functions
\begin{align}
    \pdf_\mathrm{env, 1} &= \frac{1}{1+\exp[-(C  (\rho - \rho_1))]} \quad \mathrm{and}\\
    \pdf_\mathrm{env, 2} &= \frac{1}{1+\exp[-(C  (\rho - \rho_2))]} \times \frac{1}{1+\exp[-(C  (a - a_2))]},
\end{align}
where $C$ is a compression factor characterizing the steepness of the sigmoid function.
\begin{figure}[ht!]
    \script{bayes_rho-a.py}
    \begin{centering}
        \includegraphics[width=\linewidth]{figures/analytic/Penv.pdf}
        \caption{
        Probability of abiogenesis for the OoL hypotheses $H_1$ and $H_2$ as a function of planet bulk density and semi-major axis.
        $H_1$ only requires large enough densities to exclude deep global water oceans.
            $H_2$ requires a higher minimum density due to the exposed land requirement, and small semi-major axes are excluded to prevent tidal locking.
        }
        \label{fig:Penv}
    \end{centering}
\end{figure}
Figure~\ref{fig:Penv} shows the corresponding $\pdf_\mathrm{env}$ factors and where their regions of high probability overlap.
...
